
---
title: "ldblock package: linkage disequilibrium data structures"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "February 2015"
output:
  BiocStyle::pdf_document:
    toc: yes
    number_sections: yes
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{LD block import and manipulation}
-->

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

# Introduction

There is a nice vignette in `r Biocpkg("snpStats")` concerning
linkage disequilibrium (LD) analysis as supported by software
in that package.
The purpose of this package is to simplify handling of
existing population-level data on LD for the purpose
of flexibly defining LD blocks.

# Import of HapMap LD data

The `hmld` function imports gzipped tabular data
from hapmap's repository
\url{hapmap.ncbi.nlm.nih.gov/downloads/ld_data/2009-02_phaseIII_r2/}.

```{r lkd, cache=TRUE}
suppressPackageStartupMessages(library(ldblock))
path = dir(system.file("hapmap", package="ldblock"), full=TRUE)
ceu17 = hmld(path, poptag="CEU", chrom="chr17")
ceu17
```

# A view of the block structure

For some reason knitr/render will not display this image nicely.
```{r abc, fig=TRUE, fig.width=7, fig.height=7}
library(Matrix)
image(ceu17@ldmat[1:400,1:400], 
   col.reg=heat.colors(120), colorkey=TRUE, useRaster=TRUE)
```

This ignores physical distance and MAF.  The bright stripes are
probably due to SNP with low MAF.

# Collecting SNPs exhibiting linkage to selected SNP

We'll use `ceu17` and the `gwascat` package to enumerate
SNP that are in LD with GWAS hits.

```{r getg}
library(gwascat)
data(ebicat37)
seqlevelsStyle(ebicat37) = "NCBI"
e17 = ebicat37[ which(as.character(seqnames(ebicat37)) == "17") ]
```

Some dbSNP names for GWAS hits on chr17 are
```{r getrs}
rsh17 = unique(e17$SNPS)
head(rsh17)
```

We will use `expandSnpSet` to obtain names for SNP
that were found in HapMap CEU to have which $D' > .9$
with any of these hits.  These names are added to
the input set.

```{r doexpa}
length(rsh17)
exset = expandSnpSet( rsh17, ldstruct= ceu17, lb=.9 )
length(exset)
all(rsh17 %in% exset)
```

Not all GWAS SNP are in the
HapMap LD resource.  You can use your own LD data
as long as the format agrees with that of the HapMap distribution.
